Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-16 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

\begin{thm}\label{thm:QRVPM2QIVPM-3}In a Hilbert space of dimension~$d\le3$,
if $\iota:\left[0,1\right]\rightarrow\mathscr{I}$ is subject to the
following constraints:\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations}and satisfies that for any numbers~$x_{0}$ and
$x_{1}\in\left[0,1\right]$, their sum~$x_{0}+x_{1}\in\left[0,1\right]$
implies
\begin{equation}
\iota\left(x_{0}+x_{1}\right)\subseteq\iota\left(x_{0}\right)+\iota\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
Then, $\iota\circ\mu:\events\rightarrow\mathscr{I}$ is always a QIVPM
for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$.\end{thm}

@@

\begin{thm}\label{thm:iota-statements}The following two statements
are equivalent:
\begin{enumerate}
\item \label{enu:iota-subject-to}The function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
satisfies Eqs.~(\ref{eq:iota-constraints}) and for any three numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$, their sum~$x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
implies
\begin{equation}
\begin{aligned} & \iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\\
\subseteq{} & \iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.
\end{aligned}
\label{eq:iota-Inclusion}
\end{equation}
\item \label{enu:iota-mu-QIVPM}The composited function~$\iota\circ\mu:\events\rightarrow\mathscr{I}$
is always a QIVPM for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
on any Hilbert space~$\Hilb$.
\end{enumerate}
\end{thm}

It is clear that Statement~\ref{enu:iota-subject-to} implies Statement~\ref{enu:iota-mu-QIVPM},
and the rest of the proof is to prove that if Statement~\ref{enu:iota-subject-to}
is false, then Statement~\ref{enu:iota-mu-QIVPM} is false as well.
If $\iota$ doesn't satisfy Eq.~(\ref{eq:iota-Inclusion}), we can
find three particular numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$
such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but $\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)$.
Consider the state $\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}$.
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray*}
 &  & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{eqnarray*}
and cannot be a QIVPM. Therefore, Statement~\ref{enu:iota-subject-to}
is true if and only if Statement~\ref{enu:iota-mu-QIVPM} is true.

@@

However, if we consider general QIVPMs mapping to an \emph{arbitrary}
set of intervals, there is a QIVPM that no state can be consistent
with it.

@@

\begin{definition}Assume a collection of intervals $\mathscr{I}$
including $\imposs$ and $\necess$. \end{definition}

In contrast with the QIVPM defined in Table~\ref{tab:non-Born-QIVPM},
if a QIVPM~$\iota\circ\mu_{\rho}^{\mathrm{B}}$ is induced by a conventional
quantum probability measure~$\mu_{\rho}^{\mathrm{B}}$ and a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
considered in Sec.~\ref{sec:Born}, there is always a state consistent
with it because $\rho\in\coreBorn\left(\iota\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.
Moreover, the collection of states consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to $\rho$ as the intervals get narrower and
narrower. To state this result, we introduce a few technical notions.
First, the norm of a pure state~$\ket{\psi}$ is defined as usual
$||\psi||=\sqrt{\ip{\psi}{\psi}}$. Then, the norm of the difference
of two states~$\rho$ and $\rho'$ is just the usual operator norm
\begin{equation}
\left\Vert \rho-\rho'\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert \left(\rho-\rho'\right)\ket{\psi}\right\Vert \,,
\end{equation}
which is also commonly known as $2$-norm or spectral norm~\citep{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.

>>>>

\begin{thm}\label{thm:QRVPM2QIVPM-3}In  for any quantum probability
measure~$\mu:\events\rightarrow\left[0,1\right]$.\end{thm}

@@

@@

\noindent However, if we consider general QIVPMs mapping to an \emph{arbitrary}
set of intervals, there is a QIVPM that no state can be consistent
with it.

@@

\begin{definition}Given a collection of intervals $\mathscr{I}$
including $\imposs$ and $\necess$, a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is called a \emph{morphism}\yutsung{I may need to know the precise
conditions when $\iota$ is a morphism in the sense of category theory,
or we need to find a better name for this function.}if $\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is subject to the following constraints:\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations}and satisfies that for any three numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$, their sum~$x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
implies
\begin{equation}
\begin{aligned} & \iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\\
\subseteq{} & \iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.
\end{aligned}
\label{eq:iota-Inclusion}
\end{equation}
\end{definition}

\noindent Then, $\iota$ is a morphism precisely characterizes that
$\iota\circ\mu$ is always a IVPM no matter in classical or quantum
cases.

\begin{thm}\label{thm:iota-statements}The following three statements
are equivalent:
\begin{enumerate}
\item \label{enu:iota-subject-to}The function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is a morphism.
\item \label{enu:iota-mu-CIVPM}The composited function~$\iota\circ\mu:\events\rightarrow\mathscr{I}$
is always a classical IVPM for any classical probability measure~$\mu:\events_{C}\rightarrow\left[0,1\right]$.
\item \label{enu:iota-mu-QIVPM}The composited function~$\iota\circ\mu:\events\rightarrow\mathscr{I}$
is always a QIVPM for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
on any Hilbert space~$\Hilb$.
\end{enumerate}
\end{thm}

Because the similarity between classical and quantum cases, we will
only prove the equivalence between Statement~\ref{enu:iota-subject-to}
and Statement~\ref{enu:iota-mu-QIVPM}. It is clear that $\iota$
is a morphism implies $\iota\circ\mu$ is always a QIVPM, and the
rest of the proof is to prove that if $\iota$ is not a morphism,
then $\iota\circ\mu$ might not be a QIVPM. If $\iota$ doesn't satisfy
Eq.~(\ref{eq:iota-Inclusion}), we can find three particular numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$ such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$,
but $\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)$.
Consider the state $\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}$.
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray*}
 &  & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{eqnarray*}
and cannot be a QIVPM. Therefore, $\iota$ is a morphism if and only
if $\iota\circ\mu$ is always a QIVPM.

Although the composited function~$\iota\circ\mu$ may not guarantee
to be a QIVPM when $\iota$ is not a morphism, it may still be a QIVPM
under so suitable conditions. For example, given a Hilbert space of
dimension~$d\le3$, we can guarantee $\iota\circ\mu$ is a QIVPM
when $\iota$ satisfies Eqs.~(\ref{eq:iota-constraints}) and a weaker
condition: for any numbers~$x_{0}$ and $x_{1}\in\left[0,1\right]$,
their sum~$x_{0}+x_{1}\in\left[0,1\right]$ implies
\begin{equation}
\iota\left(x_{0}+x_{1}\right)\subseteq\iota\left(x_{0}\right)+\iota\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}

In contrast with the QIVPM defined in Table~\ref{tab:non-Born-QIVPM},
if a QIVPM~$\iota\circ\mu_{\rho}^{\mathrm{B}}$ is induced by a conventional
quantum probability measure~$\mu_{\rho}^{\mathrm{B}}$ and a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
, there is always a state consistent with it because $\rho\in\coreBorn\left(\iota\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.
Moreover, the collection of states consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to $\rho$ as the intervals get narrower and
narrower. To state this result, we introduce a few technical notions.
First, the norm of a pure state~$\ket{\psi}$ is defined as usual
$||\psi||=\sqrt{\ip{\psi}{\psi}}$. Then, the norm of the difference
of two states~$\rho$ and $\rho'$ is just the usual operator norm
\begin{equation}
\left\Vert \rho-\rho'\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert \left(\rho-\rho'\right)\ket{\psi}\right\Vert \,,
\end{equation}
which is also commonly known as $2$-norm or spectral norm~\citep{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
